---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

# Resources

Here are some links I found helpful:

- https://www.computerworld.com/article/3175623/mapping-in-r-just-got-a-whole-lot-easier.html
- https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html
- http://zevross.com/blog/2018/10/02/creating-beautiful-demographic-maps-in-r-with-the-tidycensus-and-tmap-packages/
- https://geocompr.robinlovelace.net/adv-map.html#introduction-5

# Imports

These are the libraries I'm using (I silence masking warnings):

```{r Imports, message=TRUE, warning=TRUE, include=FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
library(sf)
library(raster)
library(tmap)
library(tmaptools)
library(spData)
library(spDataLarge)
library(RColorBrewer)
library(reshape2)
```

Show the current software versions I'm working with. 

```{r Session Info}
sessionInfo()
```

# Load Raw Data

This data was available from: https://www.kaggle.com/mikejohnsonjr/united-states-crime-rates-by-county/data with the following license: CC0 1.0 Universal (CC0 1.0)

```{r Load Data}
df <- read.csv("crime_data.csv")
head(df)
```

# Data Preparation

Notice how the state name is part of the "county name" variable? I'll create a new column for the state name.

```{r Extarct State Name}
# note that we trim whitespace and explicitly convert
# to a factor
df$state_abbr <- as.factor(trimws(sub('.*,\\s*', '', df$county_name)))
head(df)
```

But this gives us abbreviations, we really want the names. There's a file in this directory with the mapping.

```{r Import State Names}
states <- read.csv("state.csv")
# note that we trim whitespace and explicitly convert
# to a factor
states$state_abbr <- as.factor(trimws(states$state_abbr))
head(states)
```

We'll add the state names using a right join:

```{r Add State Names}
df <- merge(x = df, y = states, by = "state_abbr", all.x = TRUE)
head(df)
```

Next we need some aggregates by state. I'll just take the sum here, but we could choose different functions if we wanted to.

```{r Aggregate}
data <- df %>%
          dplyr::select(state_name, population, ARSON, ROBBERY, MURDER) %>%
          group_by(state_name) %>%
          summarise_all(sum)

head(data)
```



# Maps!

First we append our data into the map object. Note the warning that we only have data for the lower 48.

```{r Build the Map}
map_obj <- append_data(us_states, 
                     data,
                     key.shp = "NAME", 
                     key.data = "state_name")
```

Now we can plot this bad  boi!

```{r Plot 1}
MURDER <- tm_shape(map_obj, projection = 2163) + 
  tm_polygons("MURDER",  
              title = "MURDER", 
                id = "NAME",
                legend.hist = TRUE,
                palette = "PRGn") +
    tm_style("col_blind", legend.outside=TRUE) + 
    tm_layout(frame = FALSE)

MURDER
```

